Wood density and the growth-dependent & independent mortality

### 2015-09-02 ms.Rmd
Fixed minor typos that caused pdf to fail

---
title: "Wood density and the growth-dependent & independent mortality"
author: "J S Camac"
date: "17 August 2015"
output: pdf_document
csl: downloads/style.csl
bibliography: data/refs.bib
---

@@

where, $\mu_{dbh_{t-1}}$ defines the log of the grand mean initial dbh, and $\sigma_{dbh_{t-1}}$ defines that variation around that mean. $\mu_{dbh_{t-1}}$ was estimated using a normal prior $\textrm{N}(0, 2.5)$, which represents our prior expectation that the mean initial dbh would likely occur between 0 and 148 cm. $\sigma_{dbh{t-1}}$ was estimated using a weakly informative half cauchy prior centered on zero and a scale parameter of 2.5.

@@

\begin{equation} \label{eq:obs dbh increment}
R_i \sim \textrm{N+}(0, 5). \end{equation}.

For computational efficiency, we used the posterior means of $\dbh_t-1$, $dbh_t$ to calculate:

@@

To account for species strategy and partition variance into explained and unexplained at the observation and species level we modelled $\alpha$, $\beta$ and $\gamma$ via the following submodels:
\begin{equation} \label{eq:submodels}
\alpha_s = \alpha_{0,s} \, \rho_s^{\alpha_{1}},
\beta_s = \beta_{0,s} \, \rho_s^{\alpha_{1}},
\gamma_s = \gamma_{0,s} \, \rho_s^{\alpha_{1}}.
\end{equation}
Here $\alpha_{0,s}$,$\beta_{0,s}$,$\gamma_{0,s}$ represents the species-level random effect of species $s$ within which individual $i$ belongs and $\alpha_{1}$,$\beta_{1}$,$\gamma_{1}$ represents the effect of standardised wood density $\rho_s$ on $\alpha$,$\beta$ and $\gamma$, respectively. The form of submodels in eq \ref{eq:submodels} ensures that parameters remain positive. On a log scale these submodels \ref{eq:submodels} become additive (e.g. $\log \alpha_s = \log \alpha_{0,s} + \alpha_{1} \, \log \rho_s$, equivalent to a classic proportional hazards model widely used in survival analysis. We standardised wood density via the transformation: $\rho_s = \rho/0.6$), an operation which is equivalent to centering a log-transformed variable at a density of 0.6 g cm${^-3}$. This was done for easier interpretation, allowing $\alpha_{1,t}$ to be interpreted as the hazard for a species with wood density of 0.6cm${^-3}$ (as opposed to species responses at $\rho$ = 0; an biologically unrealistic state for a woody plant) [@Gelman:2007te].

@@

EXAMPLE:
```{r}
# white ash
median(c(0.999,0.55,0.998))
# sugar maple
median(c(0.998, 0.148, 0.999)
# American beech
median(c(0.014,0.12))
# Yellow Birch
median(c(0.555,0.995))
# Red Maple
0.99
# Eastern hemlock
0.077
# Black cherry
0.998
# White oak
0.82
# Red oak
0.985
#White pine
0.268

@@

Similarly, we used prior information from Korning and Balslev et al (1994) to estimate, $\mu_{\log \gamma_0$, the average log species effect on $\gamma$, such that:

\begin{equation} \label{eq:gamma0}
\mu_{\log \gamma_0 \sim \text{normal}\left(-3.86, 0.2\right).\\
\end{equation}

This distribution was estimated the same way as in $\mu_{\log \alpha_0$. That is we converted annual mortality probabilities of 17 tropical lowland tree species in Amazonian Ecuador that encompass the range of forest structure (i.e. understory species, small stemmed midcanopy species, large stemmed midcanopy species and canopy species) to instantaneous hazard rate, and then log transformed these and calculated the mean (i.e. -3.86) and standard error (i.e. 0.2) These annual mortality estimates were averaged over size and growth rates, and as such are likely to be an overestimate of the true value of $\gamma$. However, based on this prior, we assume that the average growth-independent species mortality will be 2.1%/yr with 95% of species growth-independent probabilities of occuring between 1.4% and 3.1%.

@@

# standard error
sd(log(k))/sqrt(length(M)) 
```

As we fitted our model to two growth measures (i.e. dbh growth and basal area growth), we used weak normal priors on $\mu_{\log \beta_0$, the average log species decay rate where:

\begin{equation} \label{eq:beta_0}
\mu_{\log \beta_0 \sim \text{normal}\left(0, 5\right).\\
\end{equation}

@@

\begin{equation} \label{eq:a2b2c2}
\alpha_2, \beta_2, \gamma_2 \sim \text{normal}\left(0, 5\right).\\
\end{equation}

@@

\begin{equation} \label{eq:a1b1c1,s}
\sigma_{a0},\sigma_{b0},\sigma_{c0} \sim \text{cauchy+}\left(0, 2.5\right).\\
\end{equation}

>>>>

---
title: "Wood density and the growth-dependent & independent mortality"
author: "J S Camac"
date: "17 August 2015"
output: pdf_document
#csl: downloads/style.csl
#bibliography: data/refs.bib
---

@@

where, $\mu_{dbh_{t-1}}$ defines the log of the grand mean initial dbh, and $\sigma_{dbh_{t-1}}$ defines that variation around that mean. $\mu_{dbh_{t-1}}$ was estimated using a normal prior $\text{N}(0, 2.5)$, which represents our prior expectation that the mean initial dbh would likely occur between 0 and 148 cm. $\sigma_{dbh{t-1}}$ was estimated using a weakly informative half cauchy prior centered on zero and a scale parameter of 2.5.

@@

\begin{equation} \label{eq:obs dbh increment}
R_i \sim \textrm{N+}(0, 5). 
\end{equation}

For computational efficiency, we used the posterior means of $dbh_{t-1}$, $dbh_t$ to calculate:

@@

To account for species strategy and partition variance into explained and unexplained at the observation and species level we modelled $\alpha$, $\beta$ and $\gamma$ via the following submodels:
\begin{equation} \label{eq:submodels}
\alpha_s = \alpha_{0,s} \, \rho_s^{\alpha_{1}},\\
\beta_s = \beta_{0,s} \, \rho_s^{\alpha_{1}}, \\
\gamma_s = \gamma_{0,s} \, \rho_s^{\alpha_{1}}. 
\end{equation}

Here $\alpha_{0,s}$,$\beta_{0,s}$,$\gamma_{0,s}$ represents the species-level random effect of species $s$ within which individual $i$ belongs and $\alpha_{1}$,$\beta_{1}$,$\gamma_{1}$ represents the effect of standardised wood density $\rho_s$ on $\alpha$,$\beta$ and $\gamma$, respectively. The form of submodels in eq \ref{eq:submodels} ensures that parameters remain positive. On a log scale these submodels \ref{eq:submodels} become additive (e.g. $\log \alpha_s = \log \alpha_{0,s} + \alpha_{1} \, \log \rho_s$, equivalent to a classic proportional hazards model widely used in survival analysis. We standardised wood density via the transformation: $\rho_s = \frac{\rho}{0.6}$), an operation which is equivalent to centering a log-transformed variable at a density of 0.6 g cm${^-3}$. This was done for easier interpretation, allowing $\alpha_{1,t}$ to be interpreted as the hazard for a species with wood density of 0.6cm${^-3}$ (as opposed to species responses at $\rho$ = 0; an biologically unrealistic state for a woody plant) [@Gelman:2007te].

@@

```{r}
# white ash
median(c(0.999,0.55,0.998))
# sugar maple
median(c(0.998, 0.148, 0.999))
# American beech
median(c(0.014,0.12))
# Yellow Birch
median(c(0.555,0.995))
# Red Maple
0.99
# Eastern hemlock
0.077
# Black cherry
0.998
# White oak
0.82
# Red oak
0.985
#White pine
0.268

@@

Similarly, we used prior information from Korning and Balslev et al (1994) to estimate, $\mu_{\log \gamma_0}$, the average log species effect on $\gamma$, such that:

\begin{equation} \label{eq:gamma0}
\mu_{\log \gamma_0} \sim \textrm{N}\left(-3.86, 0.2\right).\\
\end{equation}

This distribution was estimated the same way as in $\mu_{\log \alpha_0}$. That is we converted annual mortality probabilities of 17 tropical lowland tree species in Amazonian Ecuador that encompass the range of forest structure (i.e. understory species, small stemmed midcanopy species, large stemmed midcanopy species and canopy species) to instantaneous hazard rate, and then log transformed these and calculated the mean (i.e. -3.86) and standard error (i.e. 0.2) These annual mortality estimates were averaged over size and growth rates, and as such are likely to be an overestimate of the true value of $\gamma$. However, based on this prior, we assume that the average growth-independent species mortality will be 2.1%/yr with 95% of species growth-independent probabilities of occuring between 1.4% and 3.1%.

@@

# standard error
sd(log(k))/sqrt(length(M))

```
As we fitted our model to two growth measures (i.e. dbh growth and basal area growth), we used weak normal priors on $\mu_{\log \beta_0}$, the average log species decay rate where:

\begin{equation} \label{eq:beta_0}
\mu_{\log \beta_0} \sim \text{N}\left(0, 5\right).\\
\end{equation}

@@

\begin{equation} \label{eq:a1b1c1}
\alpha_1, \beta_1, \gamma_1 \sim \text{N}\left(0, 5\right).\\
\end{equation}

@@

\begin{equation} \label{eq:sigma}
\sigma_{a0},\sigma_{b0},\sigma_{c0} \sim \text{cauchy+}\left(0, 2.5\right).\\
\end{equation}
